### 2020-11-24 systemID_NNs.tex
making progress in writing.

\documentclass{article}
\usepackage[total={4.5in, 8in}]{geometry}
%% The graphicx and color packages should give you all the color, % figure
%inclusion, and text/figure resizing capabilities you need.
\usepackage{graphicx}
\usepackage[dvipsnames]{xcolor}
\usepackage{tikz,tikzsettings,bm}
\usepackage{almostfull}

\definecolor{pblue}{HTML}{1F77B4}
\definecolor{pgold}{HTML}{FF7f0E}
\definecolor{pgreen}{HTML}{2CA02C}
\definecolor{pred}{HTML}{D62728}
\definecolor{ppurple}{HTML}{9467BD}
\definecolor{ppink}{HTML}{E377C2}
\definecolor{pyellow}{HTML}{BCBD22}
\definecolor{pcyan}{HTML}{17BECF}

\usepackage{authblk}

%% You should not need more than this for fancy math.
\usepackage{amsmath}   % Extra math commands and environments from the AMS
\usepackage{amssymb}   % Special symbols from the AMS
\usepackage{amsthm}    % Enhanced theorem and proof environments from the AMS
\usepackage{latexsym}  % A few extra LaTeX symbols
\usepackage{cleveref}
\usepackage{lucidbry}
\input{stanacce}
\renewcommand{\baselinestretch}{1.0666}

%% The URL package is handy for typesetting URLs.  It does not define % an
%\email command because so many document styles already do that. % So we define
%one here that uses a typewriter font.
\usepackage{url}
%\DeclareRobustCommand{\email}{\begingroup \urlstyle{tt}\Url}
\providecommand{\email}{}
\renewcommand{\email}[1]{\texttt{#1}}

%% This provides various customized verbatim commands and % environments.  You
%probably don't need it.
\usepackage{fancyvrb}
\DefineShortVerb{\|} \VerbatimFootnotes
\DefineVerbatimEnvironment{code}{Verbatim}{%
  frame=single, framesep=1em, xleftmargin=1em, xrightmargin=1em, samepage=true,
  fontsize=\footnotesize}
\usepackage{upquote}

%% Your document may require a different bibliography style.  I've % come to
%prefer Author (YYYY) styles because it makes it easier for % a person who knows
%something about the literature to understand % what is being cited without
%having to skip to the list of % references.
\usepackage[authoryear,round,longnamesfirst]{natbib}

%% You won't normally need this definition in your documents, but it % is here
%so we can typeset the BibTeX logo correctly.
\makeatletter
\@ifundefined{BibTeX} {\def\BibTeX{{\rmfamily B\kern-.05em%
    \textsc{i\kern-.025em b}\kern-.08em%
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}}{}

\def\blfootnote{\xdef\@thefnmark{}\@footnotetext}
\makeatother

%% Allow more of the pages to be occupied by graphs.  These parameters % are
%described in section c.9.1 of the LaTeXbook.  Your document may % not benefit
%from these parameters (it could make things worse) so % you should only change
%these from the defaults if you need to.
\setcounter{topnumber}{2}              %% 2
\setcounter{bottomnumber}{1}           %% 1
\setcounter{totalnumber}{3}            %% 3
\renewcommand{\topfraction}{0.9}       %% 0.7
\renewcommand{\bottomfraction}{0.9}    %% 0.3
\renewcommand{\textfraction}{0.1}      %% 0.2
\renewcommand{\floatpagefraction}{.7}  %% 0.5
\newlength{\graphwidth}
\setlength{\graphwidth}{0.8\columnwidth}

\usepackage{subfiles}

\graphicspath{{./}{./figures/}{./figures/paper/}}

\newcommand{\Is}{\bm{\mathrm{I}}} % Resources.
\newcommand{\Js}{\bm{\mathrm{J}}} % Units.
\newcommand{\Ks}{\bm{\mathrm{K}}} % Resources.
\newcommand{\Ms}{\bm{\mathrm{M}}} % Transition catalog.
\newcommand{\Ns}{\bm{\mathrm{N}}} % Transition counter.

\newcommand{\useq}{\mathbf{u}} \newcommand{\xseq}{\mathbf{x}}
\newcommand{\bbR}{\mathbb{R}} \newcommand{\bbW}{\mathbb{W}}
\newcommand{\bbU}{\mathbb{U}} \newcommand{\bbI}{\mathbb{I}}
\newcommand{\bbX}{\mathbb{X}} \newcommand{\bbP}{\mathbb{P}}

\title{Hybrid nonlinear system identification using neural networks}
\author{Pratyush Kumar}
\date{\today}

\begin{document}

\maketitle

\section{Process modeling}
Consider a nonlinear plant evolving in continuous time as
\begin{align*}
  \dot{x}_p &= f_p(x_p, u, w) \\
  y &= h_p(x_p) + v
\end{align*}
in which $x_p \in \bbR^{n_p}$ is the plant state, $u \in \bbU$ is the
manipulated control input, $y \in \bbR^p$ is the measurement, $w$ is the process
noise, and $v$ is the measurement noise.

\subsection{Completion of the u-y model}
We assume that grey-box modeling is performed for the plant, and we wish to
augment the grey-box model with neural networks for improvement in the accuracy
of the model. From first-principles modeling, we have the following grey-box
model in continuous time
\begin{align*}
  \dot{x}_g &= f_g(x_g, u) \\
  y &= h_g(x_g)
\end{align*}

in which $x_g \in \bbR^{n_g}$ are the grey-box states, which are some subset of
the plant states (denote $x_g = Gx_p$) chosen to be modeled. The dynamics of
these grey-box states potentially have some missing terms in their ODEs. For
observable nonlinear systems, we can construct the original plant state using a
history of measurements and control inputs. Denote this history as 
\begin{align*}
  \mathbf{y}_{k-N_p:k-1} &= [y(k-N_p\Delta)', ... \ , y(k-\Delta)']' \\
  \mathbf{u}_{k-N_p:k-1} &= [u(k-N_p\Delta)', 
                               ... \ , u(k-\Delta)']'
\end{align*}

in which $N_p$ is the number of past measurements and $\Delta$ is the sample
time between two consecutive measurements. There exists a $\phi(\cdot)$ and a
number of delay time steps $N_p$, such that the plant state at a given time can
be written as $x_p(k) = \phi(\mathbf{y}_{k-N_p:k}, \mathbf{u}_{k-N_p:k-1})$. For
linear plants, this function $\phi(\cdot)$ is known analytically in terms of the
plant dynamic matrices.

We augment the grey-box model using neural networks as follows -
\begin{align} \label{eq:hybrid_cont_time}
\dot x_g &= f_g(x_g, u) + 
f_N(x_g, z, u) \notag \\
y &= h_g(x_g) 
\end{align}
in which $f_N(\cdot)$ is a feedforward neural network and $z(k) =
[\mathbf{y}'_{k-N_p:k-1}, \mathbf{u}'_{k-N_p:k-1}]'$. This model is converted to
discrete-time using the Runge-Kutta 4 method, and the vector of past controls
and measurements ($z$) is also viewed as a state. The discrete-time model is
represented as 
\begin{equation} \label{eq:hybrid_discrete_time}
  \begin{split}
    x_g^+ &= f_h(x_g, z, u)  \\
    z^+ &= f_z(z, x_g, u) \\
    y &= h_g(x_g)       
  \end{split}
\end{equation}
in which, $f_h(\cdot)$ is a obtained by discretization of
\eqref{eq:hybrid_cont_time} using the Runge-Kutta 4 method, and $f_z$ is known
analytically.

A few alternative modeling choices are:
\begin{itemize}
	\item Black-box: $y^+ = f_N(y, z, u)$, (purely black-box)
  \item Hybrid 1: $x_g^+ = f_N(x_g, z, u), \ z^+ = f_z(z, x_g, u), \ y =
                   h_g(x_g)$, (states have meaning)
  \item Hybrid 2: $x_g^+ = f_g(x_g, u), \ z^+ = f_z(z, x_g, u), \ y = h_g(x_g) +
                   h_N(z)$, (model residual)
  \item Hybrid 3: $\dot x_g = f_g(x_g, u, p(h_g(x_g), z)), \ z^+ = f_z(z, x_g,
  u), \ y = h_g(x_g)$, (use NNs to learn some parameters)
\end{itemize}

\subsection{Conversion to discrete time and training}
We note that the network $f_N(\cdot)$ is an addition to the grey-box dynamics in
continuous time in \eqref{eq:hybrid_cont_time}. The Runge-Kutta-4 (RK4) method
is used for the discretization as follows

\begin{align*}
  k_1 &= f_g(x_g, u) + 
  f_N(x_g, \mathbf{y}_{k-N_p:k-1}, \mathbf{u}_{k-N_p:k-1}, u) \\
  x_g & := x_g + (\Delta/2)k_1 \\
  k_2 &= f_g(x_g, u) + 
  f_N(x_g, \mathbf{y}_{k+0.5-N_p:k-0.5}, \mathbf{u}_{k-N_p:k-1}, u)\\
  x_g & := x_g + (\Delta/2)k_2 \\
  k_3 &= f_g(x_g, u) + 
  f_N(x_g, \mathbf{y}_{k+0.5-N_p:k-0.5}, \mathbf{u}_{k-N_p:k-1}, u)\\
  x_g & := x_g + \Delta k_3 \\
  k_4 &= f_g(x_g, u) + 
  f_N(x_g, \mathbf{y}_{k+1-N_p:k}, \mathbf{u}_{k-N_p:k-1}, u)\\
  x_g(k+\Delta) &= x_g(k) + (\Delta/6)(k_1 + 2k_2 + 2k_3 + k_4) \\
  \hat{y}(k+\Delta) &= h_g(x_g(k+\Delta)) 
\end{align*}
in which the sequence $\mathbf{y}_{k+0.5-N_p:k-0.5}$ is obtained by linear
interpolation.

The following optimization problem is solved to compute the weights in the
neural network
\begin{align*}
\underset{W_i, b_i}{\textnormal{min}} \sum_{k=0}^{N_s-1} 
\dfrac{1}{N_s}(\hat{y}(k) - y(k))^2
 \end{align*}
in which $W_i$, $b_i$ are the weights in each hidden layer of the NN, $N_s$ is
the number of training samples, $\hat{y}$ are the predictions of the augmented
grey-box + neural network model, $N_s$ is the number of training samples, and
$y$ are the measurements from the plant. Note that this optimization requires an
estimate of the initial grey-box state. We assume that before the open-loop
system identification experiment performed for data collection, the plant is
being operated at some steady-state such that input to output responses of the
plant and grey-box model are aligned with the use of an integrating disturbance
model. We can use the grey-box state at that steady state as the initial
condition for the above multi-step-ahead prediction-error optimization.

\subsection{Nonlinear reactor example}
We consider a nonlinear reactor with two reactions ($A \rightarrow B $ and  $3B
\rightleftharpoons C$), simulated using the following ordinary differential
equations (ODEs)
\begin{align*}
  \dfrac{dC_A}{dt} &= \dfrac{C_{Af} - C_A}{\tau} - k_1C_A\\
  \dfrac{dC_B}{dt} &= k_1C_A - 3k_2C^3_B + 3k_3C_C- \dfrac{C_B}{\tau}\\
  \dfrac{dC_C}{dt} &= k_2C^3_B - k_3C_C - \dfrac{C_C}{\tau}
\end{align*}

The grey-box model is chosen as follows with an assumption that knowledge of the
second side reaction ($3B \rightleftharpoons C $) is not available
\begin{align*}
  \dfrac{dC_A}{dt} &= \dfrac{C_{Af} - C_A}{\tau} - k_1C_A\\
  \dfrac{dC_B}{dt} &= k_1C_A - \dfrac{C_B}{\tau}
\end{align*}
The control input is $u = C_{Af}$, the states in the plant model are $x = [C_A,
C_B, C_C]'$, and the measurements are $y = [C_A, C_B]'$. We assume the following
parameters values, $k_1 = 1 \ \textnormal{m}^3/\textnormal{min}, \ k_2 = 0.01 \
\textnormal{m}^3/\textnormal{min}, \ k_3 = 0.05 \
\textnormal{m}^3/\textnormal{min}$, and $\tau = 5 \ \textnormal{min}$.

We examine the performance of the hybrid modeling framework
\eqref{eq:hybrid_discrete_time} in simulation studies. First, we generate
training data by simulating the plant model using a pseudo-random-binary-signal
(PRBS) manipulated input profile. Second, we train a black-box neural network
and the hybrid model \eqref{eq:hybrid_discrete_time} with network architectures
of [9, 16, 2] and [8, 16, 2]. Third, we examine the trained model predictions
and a mean squared error (MSE) performance metric on validation data generated
using the plant model. Finally, we analyze the steady-state cost curves obtained
using the trained models and their differences with the steady-state cost curve
of the plant model. These simulations and analyses shed light on the data
requirements of the hybrid and black-box modeling frameworks, and enable an
understanding of their performance if used for real-time economic optimization
in chemical plants.

The Figure \ref{fig:validation_pred} shows predictions of the trained black-box
and hybrid models after training with 8 hours (480 samples) of data. We observe
that both the models can match the plant data accurately. The black-box model
shows slightly worse predictions in a few places. However, both the trained
models are adequate for real-time optimization in a feedback control system.

\begin{figure}[!h]
  \centering
  \includegraphics[page=2, width=\textwidth]{tworeac_plots_nonlin.pdf}
  \caption{Model validation for the nonlinear reactor model. Both the black-box
           and hybrid model can predict the plant responses accurately. The
           model predictions shown in this plot correspond to training with 8
           hours of data.}
  \label{fig:validation_pred}
\end{figure}

We examine the prediction accuracy of the models as a function of the number of
training samples using the mean squared error (MSE) metric computed as follows
on the validation data
\begin{align*}
  \textnormal{MSE} =  \sum_{k=0}^{N_s-1} 
  \dfrac{1}{N_s}(\hat{y}(k) - y(k))^2
\end{align*}
in which $\hat{y}$ are the model predictions and $y$ denotes the plant
measurements. The Figure \ref{fig:validation_mse} shows the variation of this
MSE metric as a function of the number of training data obtained with both the
black-box and hybrid models. We observe that the hybrid model provides better
performance than the black-box model in terms of overall prediction accuracy and
data requirements.

\begin{figure}[!h]
  \centering
  \includegraphics[page=9, width=\textwidth]{tworeac_plots_nonlin.pdf}
  \caption{Mean squared error (MSE) performance metric obtained with the
           black-box and hybrid model with increase in the number of training
           samples. We notice that the hybrid model slightly outperforms the
           black-box model in data requirements.}
  \label{fig:validation_mse}
\end{figure}

Next, we study the performance of the trained models in a steady-state economic
optimization problem, and compare the cost curves obtained with the trained
models with the cost curve of the plant model. The motivation is to understand
the performance of these models if used in real-time economic optimization
layers common in chemical process industries. The optimization problem solved is

\begin{align*}
  \underset{u_s}{\textnormal{min}} & \quad \ell(y_s, u_s) \\
  x_s &= f(x_s, u_s)\\ 
  y_s &= h(x_s)
\end{align*}
in which $x_s$, $y_s$, $u_s$ are the state, output, and control respectively at
steady-states. For the nonlinear reactor example, the cost is chosen as
$\ell(y_s, u_s) = c_1C_{Af} - c_2C_B$, which is a measure of the raw material
cost and product profit. The Figure  
\ref{fig:cost_curves}

\begin{figure}[!h]
  \centering
  \includegraphics[page=6, width=\textwidth]{tworeac_plots_nonlin.pdf}
  \caption{Steady-state cost curve for the nonlinear plant, grey-box, black-box,
  residual, and hybrid model after training with 5 hours of data.}
  \label{fig:cost_curves}
\end{figure}

\begin{figure}[!h]
  \centering
  \includegraphics[page=4, width=\textwidth]{tworeac_plots_nonlin.pdf}
  \caption{Steady-state cost curve for the nonlinear plant, grey-box, black-box,
  residual, and hybrid model after training with 10 hours of data.}
  \label{fig:cost_nonlinear_10hours}
\end{figure}

\section{Conclusion}
Hey.
%\begin{figure}[!h] \centering \includegraphics[page=1,
%  width=\textwidth]{tworeac_distmodels_nonlin.pdf} \caption{Steady state
%  grey-box states and disturbance estimates based on output and input
%  disturbance models} \label{fig:cost_nonlinear_10hours} \end{figure}

\end{document}