### 2021-02-23 systemID_NNs.tex
add scripts.

\documentclass{article}
\usepackage[total={4.5in, 8in}]{geometry}
%% The graphicx and color packages should give you all the color, % figure
%inclusion, and text/figure resizing capabilities you need.
\usepackage{graphicx}
\usepackage[dvipsnames]{xcolor}
\usepackage{tikz,tikzsettings,bm}
\usepackage{almostfull}

\definecolor{pblue}{HTML}{1F77B4}
\definecolor{pgold}{HTML}{FF7f0E}
\definecolor{pgreen}{HTML}{2CA02C}
\definecolor{pred}{HTML}{D62728}
\definecolor{ppurple}{HTML}{9467BD}
\definecolor{ppink}{HTML}{E377C2}
\definecolor{pyellow}{HTML}{BCBD22}
\definecolor{pcyan}{HTML}{17BECF}

\usepackage{authblk}

%% You should not need more than this for fancy math.
\usepackage{amsmath}   % Extra math commands and environments from the AMS
\usepackage{amssymb}   % Special symbols from the AMS
\usepackage{amsthm}    % Enhanced theorem and proof environments from the AMS
\usepackage{latexsym}  % A few extra LaTeX symbols
\usepackage{cleveref}
\usepackage{lucidbry}
\input{stanacce}
\renewcommand{\baselinestretch}{1.0666}

%% The URL package is handy for typesetting URLs.  It does not define % an
%\email command because so many document styles already do that. % So we define
%one here that uses a typewriter font.
\usepackage{url}
%\DeclareRobustCommand{\email}{\begingroup \urlstyle{tt}\Url}
\providecommand{\email}{}
\renewcommand{\email}[1]{\texttt{#1}}

%% This provides various customized verbatim commands and % environments.  You
%probably don't need it.
\usepackage{fancyvrb}
\DefineShortVerb{\|} \VerbatimFootnotes
\DefineVerbatimEnvironment{code}{Verbatim}{%
  frame=single, framesep=1em, xleftmargin=1em, xrightmargin=1em, samepage=true,
  fontsize=\footnotesize}
\usepackage{upquote}

%% Your document may require a different bibliography style.  I've % come to
%prefer Author (YYYY) styles because it makes it easier for % a person who knows
%something about the literature to understand % what is being cited without
%having to skip to the list of % references.
\usepackage[authoryear,round,longnamesfirst]{natbib}

%% You won't normally need this definition in your documents, but it % is here
%so we can typeset the BibTeX logo correctly.
\makeatletter
\@ifundefined{BibTeX} {\def\BibTeX{{\rmfamily B\kern-.05em%
    \textsc{i\kern-.025em b}\kern-.08em%
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}}{}

\def\blfootnote{\xdef\@thefnmark{}\@footnotetext}
\makeatother

%% Allow more of the pages to be occupied by graphs.  These parameters % are
%described in section c.9.1 of the LaTeXbook.  Your document may % not benefit
%from these parameters (it could make things worse) so % you should only change
%these from the defaults if you need to.
\setcounter{topnumber}{2}              %% 2
\setcounter{bottomnumber}{1}           %% 1
\setcounter{totalnumber}{3}            %% 3
\renewcommand{\topfraction}{0.9}       %% 0.7
\renewcommand{\bottomfraction}{0.9}    %% 0.3
\renewcommand{\textfraction}{0.1}      %% 0.2
\renewcommand{\floatpagefraction}{.7}  %% 0.5
\newlength{\graphwidth}
\setlength{\graphwidth}{0.8\columnwidth}

\usepackage{subfiles}

\graphicspath{{./}{./figures/}{./figures/paper/}}

\newcommand{\Is}{\bm{\mathrm{I}}} % Resources.
\newcommand{\Js}{\bm{\mathrm{J}}} % Units.
\newcommand{\Ks}{\bm{\mathrm{K}}} % Resources.
\newcommand{\Ms}{\bm{\mathrm{M}}} % Transition catalog.
\newcommand{\Ns}{\bm{\mathrm{N}}} % Transition counter.

\newcommand{\useq}{\mathbf{u}} \newcommand{\xseq}{\mathbf{x}}
\newcommand{\bbR}{\mathbb{R}} \newcommand{\bbW}{\mathbb{W}}
\newcommand{\bbU}{\mathbb{U}} \newcommand{\bbI}{\mathbb{I}}
\newcommand{\bbX}{\mathbb{X}} \newcommand{\bbP}{\mathbb{P}}

\title{Hybrid nonlinear system identification using neural networks}
\author{Pratyush Kumar}
\date{\today}

\begin{document}

\maketitle

\section{Process modeling}
Consider a nonlinear plant evolving in continuous time as
\begin{align*}
  \dot{x}_p &= f_p(x_p, u, w) \\
  y &= h_p(x_p) + v
\end{align*}
in which $x_p \in \bbR^{n_p}$ is the plant state, $u \in \bbU$ is the
manipulated control input, $y \in \bbR^p$ is the measurement, $w$ is the process
noise, and $v$ is the measurement noise.

\subsection{Completion of the u-y model}
We assume that grey-box modeling is performed for the plant, and we wish to
augment the grey-box model with neural networks for improvement in the accuracy
of the model. From first-principles modeling, we have the following grey-box
model in continuous time
\begin{align*}
  \dot{x}_g &= f_g(x_g, u) \\
  y &= h_g(x_g)
\end{align*}

in which $x_g \in \bbR^{n_g}$ are the grey-box states, which are some subset of
the plant states (denote $x_g = Gx_p$) chosen to be modeled. The dynamics of
these grey-box states potentially have some missing terms in their ODEs. For
observable nonlinear systems, we can construct the original plant state using a
history of measurements and control inputs. Denote this history as 
\begin{align*}
  \mathbf{y}_{k-N_p:k-1} &= [y(k-N_p\Delta)', ... \ , y(k-\Delta)']' \\
  \mathbf{u}_{k-N_p:k-1} &= [u(k-N_p\Delta)', 
                               ... \ , u(k-\Delta)']'
\end{align*}

in which $N_p$ is the number of past measurements and $\Delta$ is the sample
time between two consecutive measurements. There exists a function $\phi(\cdot)$
and a number of delay time steps $N_p$, such that the plant state at a given
time can be written as $x_p(k) = \phi(\mathbf{y}_{k-N_p:k},
\mathbf{u}_{k-N_p:k-1})$. For linear plants, this function $\phi(\cdot)$ is
known analytically in terms of the plant dynamic matrices.

We augment the grey-box model using neural networks as follows -
\begin{align} \label{eq:hybrid_cont_time}
\dot x_g &= f_g(x_g, u) + 
f_N(x_g, z, u) \notag \\
y &= h_g(x_g) 
\end{align}
in which $f_N(\cdot)$ is a feedforward neural network and $z(k) =
[\mathbf{y}'_{k-N_p:k-1}, \mathbf{u}'_{k-N_p:k-1}]'$. This model is converted to
discrete-time using the Runge-Kutta 4 method, and the vector of past controls
and measurements ($z$) is also viewed as a state. The discrete-time model is
represented as 
\begin{equation} \label{eq:hybrid_discrete_time}
  \begin{split}
    x_g^+ &= f_h(x_g, z, u)  \\
    z^+ &= f_z(z, x_g, u) \\
    y &= h_g(x_g)       
  \end{split}
\end{equation}
in which, $f_h(\cdot)$ is a obtained by discretization of
\eqref{eq:hybrid_cont_time} using the Runge-Kutta 4 method, and $f_z$ is known
analytically.

A few alternative modeling choices are:
\begin{itemize}
  \item Black-box: $y^+ = f_N(y, z, u), \ z^+ = f_z(z, y, u)$
        (purely black-box).
  \item Hybrid 1: $x_g^+ = f_N(x_g, z, u), \ z^+ = f_z(z, x_g, u), \ y =
                   h_g(x_g)$, (states have meaning)
  \item Hybrid 2: $x_g^+ = f_g(x_g, u), \ z^+ = f_z(z, x_g, u), \ y = h_g(x_g) +
                   h_N(z)$, (model residual)
  \item Hybrid 3: $\dot x_g = f_g(x_g, u, p(h_g(x_g), z)), \ z^+ = f_z(z, x_g,
  u), \ y = h_g(x_g)$, (use NNs to learn some parameters)
\end{itemize}

\section{Conversion to discrete time and training}
We note that the network $f_N(\cdot)$ is an addition to the grey-box dynamics in
continuous time in \eqref{eq:hybrid_cont_time}. The Runge-Kutta-4 (RK4) method
is used for the discretization as follows

\begin{align*}
  x_{gk} &= x_g(k) \\
  k_1 &= f_g(x_g, u) + 
  f_N(x_g, \mathbf{y}_{k-N_p:k-1}, \mathbf{u}_{k-N_p:k-1}, u) \\
  x_g & := x_g + (\Delta/2)k_1 \\
  k_2 &= f_g(x_g, u) + 
  f_N(x_g, \mathbf{y}_{k+0.5-N_p:k-0.5}, \mathbf{u}_{k-N_p:k-1}, u)\\
  x_g & := x_g + (\Delta/2)k_2 \\
  k_3 &= f_g(x_g, u) + 
  f_N(x_g, \mathbf{y}_{k+0.5-N_p:k-0.5}, \mathbf{u}_{k-N_p:k-1}, u)\\
  x_g & := x_g + \Delta k_3 \\
  k_4 &= f_g(x_g, u) + 
  f_N(x_g, \mathbf{y}_{k+1-N_p:k}, \mathbf{u}_{k-N_p:k-1}, u)\\
  x_g(k+\Delta) &= x_{gk} + (\Delta/6)(k_1 + 2k_2 + 2k_3 + k_4) \\
  \hat{y}(k+\Delta) &= h_g(x_g(k+\Delta)) 
\end{align*}
in which the sequence $\mathbf{y}_{k+0.5-N_p:k-0.5}$ is obtained by linear
interpolation.

The following optimization problem is solved to compute the weights in the
neural network
\begin{align*}
\underset{W_i, b_i}{\textnormal{min}} \sum_{k=0}^{N_s-1} 
\dfrac{1}{N_s}(\hat{y}(k) - y(k))^2
\end{align*}
in which $W_i$, $b_i$ are the weights in each hidden layer of the network, $N_s$
is the number of training samples, $\hat{y}$ are the predictions of the
hybrid grey-box + neural network model, $N_s$ is the number of training
samples, and $y$ are the measurements from the plant. Note that this
optimization requires an estimate of the initial grey-box state. We assume that
before the open-loop system identification experiment is performed for data
collection, the plant is being operated at some steady-state such that input to
output responses of the plant and grey-box model can be aligned with the use of an
integrating disturbance model. We can use the grey-box state at that steady
state as the initial condition for the above multi-step-ahead prediction-error
optimization.

\section{Nonlinear reactor example}
We now demonstrate the application of the above hybrid modeling framework
via a simulation study on a simple chemical process example. We consider a nonlinear reactor with two reactions ($A \rightarrow B $ and  $3B
\rightleftharpoons C$), simulated using the following ordinary differential
equations (ODEs)
\begin{align*}
  \dfrac{dC_A}{dt} &= \dfrac{C_{Af} - C_A}{\tau} - k_1C_A\\
  \dfrac{dC_B}{dt} &= k_1C_A - 3k_2C^3_B + 3k_3C_C- \dfrac{C_B}{\tau}\\
  \dfrac{dC_C}{dt} &= k_2C^3_B - k_3C_C - \dfrac{C_C}{\tau}
\end{align*}

The grey-box model is chosen as follows with an assumption that knowledge of the
second side reaction ($3B \rightleftharpoons C $) is not available
\begin{align*}
  \dfrac{dC_A}{dt} &= \dfrac{C_{Af} - C_A}{\tau} - k_1C_A\\
  \dfrac{dC_B}{dt} &= k_1C_A - \dfrac{C_B}{\tau}
\end{align*}
The control input is $u = C_{Af}$, the states in the plant model are $x = [C_A,
C_B, C_C]'$, and the measurements are $y = [C_A, C_B]'$. We assume the following
parameter values, $k_1 = 1 \ \textnormal{m}^3/\textnormal{min}, \ k_2 = 0.01 \
\textnormal{m}^3/\textnormal{min}, \ k_3 = 0.05 \
\textnormal{m}^3/\textnormal{min}$, and $\tau = 5 \ \textnormal{min}$.

We examine the performance of the hybrid modeling framework
\eqref{eq:hybrid_discrete_time} in simulation studies. First, we generate
training data by simulating the plant model using a pseudo-random-binary-signal
(PRBS) of the manipulated control input. Second, we train a black-box neural network
and the hybrid model \eqref{eq:hybrid_discrete_time} with network architectures
of [9, 16, 2] and [8, 16, 2]. Third, we examine the trained model predictions
and a mean squared error (MSE) performance metric on validation data generated
using the plant model. Finally, we analyze the steady-state cost curves obtained
using the trained models and their differences with the steady-state cost curve
of the plant model. These simulations and analyses shed light on the data
requirements of the hybrid and black-box modeling frameworks, and enable an
understanding of their performance if used for real-time economic optimization
purposes in chemical plants.

The Figure shows predictions of the trained black-box
and hybrid models after training with 8 hours (480 samples) of data. We observe
that both the models can match the plant data accurately. The black-box model
shows slightly worse predictions in a few places. However, both the trained
models are adequate for real-time optimization purpose in a feedback control system.

We examine the prediction accuracy of the models as a function of the number of
training samples using the mean squared error (MSE) metric computed as follows
on the validation data
\begin{align*}
  \textnormal{MSE} =  \sum_{k=0}^{N_s-1} 
  \dfrac{1}{N_s}(\hat{y}(k) - y(k))^2
\end{align*}
in which $\hat{y}$ are the model predictions and $y$ denotes the plant
measurements. The Figure  shows the variation of this
MSE metric as a function of the number of training data obtained with both the
black-box and hybrid models. We observe that the hybrid model provides better
performance than the black-box model in terms of overall prediction accuracy and
data requirements.

Next, we study the performance of the trained models in a steady-state economic
optimization problem, and compare the cost curves obtained with the trained
models with the cost curve of the plant model. The motivation is to understand
the performance of these models if used in real-time economic optimization
layers common in chemical process industries. The optimization problem solved is

\begin{align*}
  \underset{u_s}{\textnormal{min}} & \quad \ell(y_s, u_s) \\
  x_s &= f(x_s, u_s)\\ 
  y_s &= h(x_s) \\
  \underline{u} \leq &u_s \leq \overline{u}
\end{align*}
in which $x_s$, $y_s$, $u_s$ are the state, output, and control respectively at
steady-states. For the nonlinear reactor example, the cost is chosen as
$\ell(y_s, u_s) = c_1C_{Af} - c_2C_B$, which measures the raw material cost and
product profit. The Figure shows the cost curves of the
plant, grey-box, black-box, and hybrid model obtained with 8 hours of training
data. We first notice from the cost curve of the grey-box model that an
optimization with the grey-box model results in the optimizer recommending a
full supply of the raw material $A$. Since some of the product $B$ reacts to
form an undesired side product $C$ in the plant, this recommendation incurs a
cost as observed from the plant cost curve. An intuitive approach to correct
this type of plant-model mismatch is to augment the grey-box model with
integrating disturbance models and estimate the plant model mismatch from data
at one steady state. This strategy however, cannot recover the plant cost curve
in the entire actuator constraint set as the integrating disturbance models are
often linear. We notice that the cost curve of the hybrid model is almost the
same as the cost curve of the plant model in the entire actuator constraint set.
The cost curve of the black-box model is also shown and we notice that it is not
as closely aligned with the plant as compared to the hybrid model.

Finally, we analyze the variation in the cost curves of the black-box and hybrid
models with increasing number of training samples. We analyze the following cost
error metric in the entire actuator constraint set
\begin{align*}
  e(u_s) = \vert \ell(y_p(u_s), u_s) - \ell(y_m(u_s), u_s) \vert
\end{align*}
in which $y_p(\cdot)$ and $y_m(\cdot)$ denote the plant and model steady state
output respectively. The Figures and show the variation in this cost error metric in the
actuator constraint set with increasing number of training samples for the
black-box and hybrid model respectively. We observe that with an increase in the
number of training samples, the error metric for the hybrid model approaches
near zero values for a wide range of actuator values in the constraint set. The
error metric of the black-box model remains relatively larger than that of the
hybrid model in the constraint set with increase in training data. The
implication of this error metric going to zero for all actuator values in the
constraint set is that the optimum computed for the steady-state economic
optimization problem with the trained model will be the same as the true plant
optimum.

\section{Large, CSTR and flash in series example}
\begin{figure}[!h]
  \centering
  \includegraphics[width=\textwidth, height=0.4\textheight]{cstr_flash.pdf}
  \caption{Schematic of the CSTR and flash plant. The CSTR facilitates two 
  reactions, the first reaction ($A \rightarrow B$) forms the primary product $B$, and the second reaction ($3B \rightarrow C$) 
  consumes the desired product $B$ and forms the undesired product $C$.
  The non-adiabatic flash separates the reactant $A$ and 
  products $B$ and $C$. The major component in the vapor phase 
  is $A$, and it is recycled back after cooling to the CSTR.}
  \label{fig:cstr_flash}
\end{figure}
%\newpage
\textbf{Plant -- }
\begin{align*}
  \dfrac{dH_r}{dt} &= \dfrac{F + D -F_r}{A_r}\\
  \dfrac{dC_{Ar}}{dt} &= \dfrac{F(C_{Af} -C_{Ar}) +
                         D(C_{Ad} -C_{Ar})}{A_rH_r} - r_1 \\
  \dfrac{dC_{Br}}{dt} &= \dfrac{-FC_{Br} + 
                          D(C_{Bd} -C_{Br})}{A_rH_r} + r_1 -3r_2\\
  \dfrac{dC_{Cr}}{dt} &= \dfrac{-FC_{Cr} + 
  D(C_{Cd} -C_{Cr})}{A_rH_r} + r_2\\
  \dfrac{dT_r}{dt} &= \dfrac{F(T_f - T_r) + D(T_d -T_r)}{A_rH_r} + 
                      \dfrac{r_1\Delta H_1 + r_2\Delta H_2}{\rho C_p} + 
                      \dfrac{Q_r}{\rho A_r C_p H_r}\\
\end{align*}

\begin{align*}
  \dfrac{dH_b}{dt} &= \dfrac{F_r - F_b - D}{A_b} \\
  \dfrac{dC_{Ab}}{dt} &= \dfrac{F_r(C_{Ar} -C_{Ab}) + 
                          D(C_{Ab} -C_{Ad})}{A_bH_b} \\
  \dfrac{dC_{Bb}}{dt} &= \dfrac{F_r(C_{Br} -C_{Bb}) + 
                          D(C_{Bb} -C_{Bd})}{A_bH_b} \\
  \dfrac{dC_{Cb}}{dt} &= \dfrac{F_r(C_{Cr} -C_{Cb}) + 
                          D(C_{Cb} -C_{Cd})}{A_bH_b} \\
  \dfrac{dT_b}{dt} &= \dfrac{F_r(T_r - T_b)}{A_bH_b} +
                      \dfrac{Q_b}{\rho A_b C_p H_b}\\
\end{align*}

The sample time for plant measurements is $1$ min, the 
plant states are ($H_r, C_{Ar}, C_{Br}, C_{Cr}, T_r, 
H_b, C_{Ab}, C_{Bb}, C_{Cb}, T_b$), and the measurements are just 
the heights and temperatures of the CSTR and flash. \\

\textbf{Grey-box model -- }
\begin{align*}
  \dfrac{dH_r}{dt} &= \dfrac{F + D -F_r}{A_r}\\
  \dfrac{dC_{Ar}}{dt} &= \dfrac{F(C_{Af} -C_{Ar}) + 
                         D(C_{Ad} -C_{Ar})}{A_rH_r} - r_1 \\
  \dfrac{dC_{Br}}{dt} &= \dfrac{-FC_{Br} + 
                          D(C_{Bd} -C_{Br})}{A_rH_r} + r_1\\
  \dfrac{dT_r}{dt} &= \dfrac{F(T_f - T_r) + D(T_d -T_r)}{A_rH_r} + 
                      \dfrac{r_1\Delta H_1}{\rho C_p} + 
                      \dfrac{Q_r}{\rho A_r C_p H_r}\\
  \dfrac{dH_b}{dt} &= \dfrac{F_r - F_b - D}{A_b} \\
  \dfrac{dC_{Ab}}{dt} &= \dfrac{F_r(C_{Ar} -C_{Ab}) + 
                          D(C_{Ab} -C_{Ad})}{A_bH_b} \\
  \dfrac{dC_{Bb}}{dt} &= \dfrac{F_r(C_{Br} -C_{Bb}) + 
                          D(C_{Bb} -C_{Bd})}{A_bH_b} \\
  \dfrac{dT_b}{dt} &= \dfrac{F_r(T_r - T_b)}{A_bH_b} +
                      \dfrac{Q_b}{\rho A_b C_p H_b}\\
\end{align*}

The grey-box model assumes no knowledge of the second 
reaction ($3B \rightarrow C$), which produces the undesired 
product $C$. Note that since only heights and temperatures are 
measured, all the states in the plant model are 
not observable and only asymptotically detectable.

The economic objective is -- 
\begin{align*}
  \underset{u_s}{\textnormal{min}} \quad 
  \ell(u_s, x_s) \quad := & \quad c_1FC_{Af} + c_2Q_r + 
                                      c_2Q_b + c_2D\rho C_p(T_d-T_b) - 
                                      c_3F_bC_{Bb} \\
  x_s &= f(x_s, u_s)\\ 
  y_s &= h(x_s) \\
  \underline{u} \leq &u_s \leq \overline{u}
\end{align*}

in which $c_1$ is the raw material cost of unit $\$$/mol-A,
$c_2$ is the energy price of unit $\$$/kW, 
and $c_3$ is the product selling price of unit $\$$/mol-B.
\end{document}