### 2021-03-10 sysID.tex
now focus on the Koopman model for the smaller example.

\documentclass[10pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{authblk}
\usepackage{amssymb}
%\usepackage{lucidbry}
\usepackage{graphicx}
\usepackage{latexsym}
\usepackage{subcaption}
\usepackage{physics}
\usepackage[dvipsnames]{xcolor}
\usepackage{tikz,tikzsettings}
\usepackage{verbatim}
\usetikzlibrary{arrows,shapes}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{verbatim}
\usepackage{float}
\usepackage[round]{natbib}
\usepackage[english]{babel}
\bibliographystyle{plainnat}
%\setcitestyle{numeric,open={(},close={)}} \usepackage{color}
%\usepackage{colortbl}
\usetikzlibrary{decorations.markings}
\usetikzlibrary{decorations.pathreplacing}
\usetikzlibrary{shapes,calc,backgrounds,arrows,fit,decorations.pathmorphing,matrix}
\usepackage[margin=0.79in]{geometry}

\graphicspath{{./}{./figures/}{./figures/paper/}}

\newcommand{\Is}{\bm{\mathrm{I}}} % Resources.
\newcommand{\Js}{\bm{\mathrm{J}}} % Units.
\newcommand{\Ks}{\bm{\mathrm{K}}} % Resources.
\newcommand{\Ms}{\bm{\mathrm{M}}} % Transition catalog.
\newcommand{\Ns}{\bm{\mathrm{N}}} % Transition counter.

\newcommand{\useq}{\mathbf{u}} \newcommand{\xseq}{\mathbf{x}}
\newcommand{\bbR}{\mathbb{R}} \newcommand{\bbW}{\mathbb{W}}
\newcommand{\bbU}{\mathbb{U}} \newcommand{\bbI}{\mathbb{I}}
\newcommand{\bbX}{\mathbb{X}} \newcommand{\bbP}{\mathbb{P}}

\title{Nonlinear system identification using machine learning tools}
\author{Pratyush Kumar}
\date{\today}

\begin{document}

\maketitle

\section{Process modeling}
We consider the problem of system identification of the following nonlinear plant.
\begin{align*}
  x_p^+ &= f_p(x_p, u, w) \\
  y &= h_p(x_p) + v
\end{align*}
in which $x_p \in \bbR^{n_p}$ is the plant state, $u \in \bbU \subset \bbR^m$ is the
manipulated control input, $y \in \bbR^p$ is the measurement, $w$ is the process
noise, and $v$ is the measurement noise. The objective of developing the model is to use it subsequently in model predictive control (MPC). Therefore, we consider several types of grey-box, hybrid, and black-box models and analyze their predictive capabilities as well as their usefulness for online optimization in MPC.

\section{Model architectures for system identification}
We describe several process modeling frameworks which determine parameters in their proposed model architecture by minimizing the multi-step ahead prediction error of the model on training data. A nonlinear optimization problem is solved for this identification step. The estimated parameters define the model, which is then used subsequently to perform model validation and for use in MPC.

\subsection{Grey-box model}
We first consider the grey-box modeling strategy, which is the current industrial state-of-the-art system identification technique. A parameterized model is defined based on first-principle knowledge about the plant, and parameters in the model are estimated from training data by solving the prediction-error minimization problem. The grey-box model is typically defined in continuous time and is of the following form:

\begin{align*}
  \dot{x}_g &= f_g(x_g, u, \theta_g) \\
  y &= h_g(x_g, \theta_g)
\end{align*}

The following nonlinear program (NLP) is solved to identify the parameters $\theta_g$:

\begin{align} \label{eq:gb_training}
  \underset{x_g^i(0), \ \theta_g}{\textnormal{min}} \sum_{i=1}^{N_{tr}} \sum_{k=1}^{N_t} 
  \dfrac{1}{N_tN_{tr}} &\norm{\tilde{y}^i(k) - y^i(k)}^2_{R_v^{-1}}  \\
  \textnormal{subject to} \quad \quad x_g^{i+} &= f_{gd}(x_g^i, u^i, \theta_g) \nonumber \\
   \tilde{y}^i &= h_g(x_g^i, \theta_g) \nonumber
\end{align}

in which $\theta_g$ and $x_g^i(0)$ are the decision variables in the optimization problem, which are parameters in the grey-box model and initial state used for forecasting the predicted measurements. We assume that multiple trajectories can be used in the training data, and the superscript $i$ is used to denote the trajectory number. We use $N_{tr}$ to denote the total number of trajectories in training data and $N_t$ to denote the number of samples in each trajectory. The function $f_{g}(\cdot)$ is discretized at the measurement sample time to obtain the function $f_{gd}(\cdot)$ that is used to make forecasts in the above optimization.

\subsection{Hybrid model}

Often in industrial applications, first-principle knowledge about the plant is incomplete. Hence, grey-box system identification and subsequent use of the grey-box model in real-time optimization can lead to suboptimal economic performance. We propose to augment the grey-box model using a neural network (NN) such that the overall hybrid model can capture any unknown process dynamics in the plant, therefore leading to improved economic performance. We define this hybrid model as follows:

\begin{align*}
  x_g^+ &= f_{gd}(x_g, u, \theta_g) + f_N(z, u, \theta_N) \\
  z^+ &= f_z(x_g, z, u) \\
  y &= h_g(x_g, \theta_g)
\end{align*}
in which $f_N(\cdot)$ is a NN and $\theta_N$ denotes the parameters in the network. We use $z$ to denote the vector of past measurements and control inputs. This vector is considered as an additional state in the overall dynamic model and is defined as follows:
\begin{align} \label{eq:pastyu}
  \mathbf{y}_{k-N_p:k-1} &= [y(k-N_p)', ... \ , y(k-1)']' \nonumber \\
  \mathbf{u}_{k-N_p:k-1} &= [u(k-N_p)', ... \ , u(k-1)']' \nonumber \\
  z(k) &= [\mathbf{y}_{k-N_p:k-1}', \mathbf{u}_{k-N_p:k-1}']'
\end{align}
in which $N_p$ is the number of past measurements and control inputs used. The function $f_z(\cdot)$ describes the dynamics of the state $z$ and its structure is defined as follows: 
\begin{align*}
  f_z(x_g, z, u) = \begin{bmatrix}
    y(k-N_p+1) \\
    y(k-N_p+2) \\
    \vdots \\
    h_g(x_g, \theta_g) \\
    u(k-N_p+1) \\ 
    u(k-N_p+2) \\
    \vdots \\
    u(k)
  \end{bmatrix}
\end{align*}
We assume that the parameters in the grey-box model are first estimated by solving the NLP \eqref{eq:gb_training}, and the following similar NLP is solved to determine parameters ($\theta_N$) in the NN.

\begin{align} \label{eq:hybrid_training}
  \underset{\theta_N}{\textnormal{min}} \sum_{i=1}^{N_{tr}} \sum_{k=1}^{N_t} 
  \dfrac{1}{N_tN_{tr}} &\norm{\tilde{y}^i(k) - y^i(k)}^2  \\
  \textnormal{subject to} \quad \quad x_g^{i+} &= f_{gd}(x_g^i, u^i, \theta_g) + f_z(z^i, u^i, \theta_N) \nonumber \\
  z^{i+} &= f_z(x_g^i, z^i, u^i) \nonumber \\
   \tilde{y}^i &= h_g(x_g^i, \theta_g) \nonumber
\end{align}
The stochastic gradient descent algorithm Adam is used to solve the above optimization using the software tensorflow. Note that we do not estimate the initial state for each trajectory, and fix those states heuristically to make forecasts during the optimization process. The control inputs and measurements are scaled as $u := (u - u_{\textnormal{MEAN}})/u_{\textnormal{STD}}$ and 
$y := (y - y_{\textnormal{MEAN}})/y_{\textnormal{STD}}$, in which $u_{\textnormal{MEAN}}$ and $y_{\textnormal{MEAN}}$ are the respective means computed on training data. Similarly, $u_{\textnormal{STD}}$ and $y_{\textnormal{STD}}$ are the respective standard deviations. Due to this scaling, we do not consider the penalty $R_v^{-1}$ in the training objective.

\subsection{Black-box models}
Next, we consider black-box system identification strategies which do not impose any a-prior knowledge about the plant and develop the dynamic model purely from data. We consider two types of black-box models in this section: NNs and dynamic models based on the Koopman operator theory. The latter type of model is considered based on the motivation that it imposes a linear structure on the dynamics in a high-dimensional space, therefore leading to a linear or quadratic program to be solved online in MPC.

\subsubsection{Neural network}
The black-box NN uses the vector of past measurements and control inputs $z$ \eqref{eq:pastyu} as the state in the dynamic model and is structured as follows
\begin{align*}
  z^+ &= f_z(z, u) \\ 
  y &= h_N(z, \theta_N)
\end{align*}
in which $h_N(\cdot)$ is a NN and $\theta_N$ are the parameters in the network. The function $f_z(\cdot)$ denotes the dynamics of the state $z$ and is defined as

\begin{align*}
  f_z(z, u) = \begin{bmatrix}
    y(k-N_p+1) \\
    y(k-N_p+2) \\
    \vdots \\
    h_N(z, \theta_N) \\
    u(k-N_p+1) \\ 
    u(k-N_p+2) \\
    \vdots \\
    u(k)
  \end{bmatrix}
\end{align*}

\subsubsection{Koopman operator model}

\subsubsection{Training}
The training problem is to find the parameters in the neural network
$f_N$ and the linear model matrices $A$ and $B$ by minimizing the following mean squared error:

\begin{align*}
  \underset{A, B, W_i, b_i}{\textnormal{min}} \sum_{k=0}^{N_t-1} 
  \dfrac{1}{N_t}(\hat{x}(k) - x(k))^2
\end{align*}

\section{Two reaction system plant model}

\section{Model validation}

\section{Steady state optimization}

\section{Dynamic economic optimization}

\section{Future work}

\end{document}